---
title: "Description of emeScheme"
subtitle: "emeScheme Version v`r read.dcf(here::here('DESCRIPTION'), all = TRUE)[['Version']]`"
author: "Rainer M Krug <Rainer@uzh.ch>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Description of emeScheme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(emeScheme)
library(here)
library(magrittr)
library(dplyr)
library(knitr)
library(kableExtra)
library(plantuml)
#
v <- read.dcf(here::here('DESCRIPTION'), all = TRUE)[['Version']]
```

# Introduction

This scheme will provide a framework for describing aquatic microcosm experiments in the sense as used in the papers in the [References] section.

It aims at providing a structured way to describe the data and to make finding the data using the provided metadata possible.

This scheme **does neither** aim at covering aspects of the actual analysis of the extracted data **nor** does it aim at giving all information to re-run the experiment.

# Terminology and Conventions

## Terminology

* **property set:** a property containing other **property sets** or **data properties**
* **data property:** a property containing only **value properties**. Equivalent to a table.
* **value property:** a column in a **data property**.
* **emeScheme:** the name of this scheme. All properties inherit from the **emeScheme class**
* **emeSchemeSet:** a property set in the emeScheme. It contains the following attributes
    * **propertyName:** the name of the properties contained in this emeSchemeSet
    * **class:** a string defining the class (**emeSchemeSet**, **emeScheme**, plus internal ones)
* **emeSchemeData:** a value property in the emeSAcheme, i.e. a table containing metadata. It contains the following attributes:
    * **propertyName:** the name of the property
    * **names:** the column names of the table
    * **units:** the units used each in the column
    * **type:** the type of each column. Can be **numeric** (== **double**), **integer**, **logical** or **character** (== empty)
    * **allowedValues:** a string of comma separated allowed values. *...* indicates that additional velues are allowed
    * **class:** a string defining the class (**emeSchemeData_PROPERTYNAME**, **emeSchemeData**, **emeScheme**, plus internal ones)

## Conventions

* Properties starting with a **capital letters** are **property sets** or **data properties**. 
* Properties starting with a **small letter** are **value properties**.

# Relationship to other schemes

The emeScheme is a metadata scheme for Ecological Microcosm Experiments. As these are essentially ecological data, the use of other schemes geareed towards ecological data also comes to mind. One widely used scheme is the [eml scheme](https://knb.ecoinformatics.org/external//emlparser/docs/index.html). 

The main difference between the emeScheme and other ecological meta data schemes is that in the development of the emeScheme the aim was to develop a scheme specific for a certain type of experiment. This specificity went together with the objective of keeping the scheme simple to fill in and to understand. This resulted in a meta data scheme which contains all information necessary to describe the data generated in these aquatic ecological microcosm experiments, while at the same time being simple enough to be filled in by the researcher without to much time required. It should be possible to use e.g. the afore mentioned [eml scheme](https://knb.ecoinformatics.org/external//emlparser/docs/index.html) to store the meta data contained in the emeScheme, but it would require a much larger investment in time to fill in the eml scheme as it is much more general per definition.

The emeScheme is neither intended nor suited for experiments outside of ecological and aquatic microcosm experiments.




# Structure of the emeScheme

The **emeScheme** is providing meta data for a **data file bundle**. A **data bundle** is an archive (e.g. tar.gz or zip) cosisting of multiple **data files** and one file with the metadata **<span style="color:red">TODO we have to decide on the format of this file - should be a text file?</span>**. If the data files represent tabular data, they should be in csv format, otherwise any open format. 

It is a **property set** (`emeSchemeSet`) which contains of five different sets of **data properties** (`emeSchemeData`) which are tables of metadata.

These five **data properties** are:

1. **Experiment**: General informtion about the experiment. Each dataset can only have one.
2. **Species**: Species used in the experiment. Each dataset can have multiple of these.
3. **Treatment**: Description of the treatment and their differences. Each dataset can have multiple of these. 
4. **Measurement**: Measurement methodology and parameter. Each dataset can have multiple of these.
5. **DataExtraction**: Methodology of extraction to extract the data to be analysed from the raw data resulting from the Measurement. Each dataset can have multiple of these.
6. **TableMetaData**: Meta data for each data file. If the data file contains tabular data, a description of the columns and a specification of the column containing the different parameters between treeatments, if other format, e.g. video, a short description of the video and the name of the treatment. 


It is important to note that

1. All **data properties** can obtain multiple rows (i.e. are repeatable) of metadata with the exception of the **Experiment** which can contain only one row of metadata.
2. Many **value properties** have an attribute `allowedValues`. If `...` is included in this attribute, any value can be entered, but, if possible, one should choose a value from the list. 

```{r data_layout, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 15, fig.height = 15}
'
!define Project(name,desc) class name as "desc" << (P,pink) >> #TECHNOLOGY
!define Data(name,desc) class name as "desc" << (D,crimson) >> #TECHNOLOGY
!define MetaData(name,desc) class name as "desc" << (M,lime) >>

Project(PhysExperiment, Experiment) {
  {field} In the Lab
}


MetaData(Experiment, Experiment) {
  {field} Experimental MetaData
  {method} + Single Experiment
}


MetaData(Species, Species) {
  {field} Species metadata
  {method} + List of Species
}


MetaData(Treatment, Treatment) {
  {field} Treatment MetaData
  {method} + List of Treatments
}



MetaData(Measurement, Measurement) {
  {field} Measurement MetaData
  {method} + List of Measurings
}

Data( RawData, Raw Data DataFile) {
  {field} Contains Raw Data
  {method} - dataCite metadata
  {method} - emeScheme metadata
  {method} + Experiment
  {method} + Species
  {method} + Set of Treatment (names)
  {method} + Set of Measurements (names)
}

MetaData( DataExtraction, DataExtraction) {
  {field} Extraction Method MetaData
  {method} + List of DataExtractions
}


Data( ExtractedData, Extracted Data DataFile) {
  {field} Contains Extracted Data
  {method} - dataCite metadata
  {method} - emeScheme metadata
  {method} - reference to Raw Data DataFile
  {method} + Experiment
  {method} + Species
  {method} + Set of Treatment (names)
  {method} + Set of Measurement (names)
  {method} + Set of DataExtraction (names)
}


PhysExperiment "1" *-- "many" Experiment : contains

Experiment "1" *-- "many" Species : contains
Experiment "1" *-- "many" Treatment : contains
Experiment "1" *-- "many" Measurement : contains
Experiment "1" *-- "many" DataExtraction : contains

Species "1" *-- "many" RawData : contains
Treatment "1" *-- "many" RawData : contains
Measurement "1" *-- "many" RawData : contains

Species "1" *-- "many" ExtractedData : contains
Treatment "1" *-- "many" ExtractedData : contains
Measurement "1" *-- "many" ExtractedData : contains
DataExtraction "1" *-- "many" ExtractedData : contains
' %>%
  plantuml %>%
  plot()
```



# The emeScheme
This document was than re-ordered and thinned which resulted in the initial verions of [the Google Doc Sheet](https://docs.google.com/spreadsheets/d/1sCImO4VadNwNKuqgtC3CPksS95Tl6SJVtY78aFo9gLg).

Here we describe the structure of the **emeScheme**

## emeScheme Structure
```{r struct_emeScheme}
print(emeScheme, printData = F, printAttr = FALSE, printExtAttr = FALSE)
```

### Experiment

```{r struct_Experiment}
emeScheme_raw$Experiment %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

### Species

```{r Species}
emeScheme_raw$Species %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

### Treatment

```{r struct_Treatment}
emeScheme_raw$Treatment %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

### Measurement

```{r struct_Measurement}
emeScheme_raw$Measurement %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

### DataExtraction

```{r struct_DataExtraction}
emeScheme_raw$DataExtraction %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

### DataFile

```{r struct_DataFile}
emeScheme_raw$DataFile %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```


A **property set** which contains **data properties** must be seen as tables and must therefore have the same number of entries for each data property.

The structure as at **`r read.dcf(here("DESCRIPTION"))[1, "GSUpdate"]` GMT** is as followed:


# Example data

An xml file with the example data can be downloaded from [here](./data/emeScheme_example.xml)

# XSD Grammar
The xsd grammar has been generated using [xmlgrid.net](http://xmlgrid.net/xml2xsd.html).
You can download it from [here - right mouse click - Save Linked Content](./data/emeScheme.xsd.xml)






























































