---
title: "Howto Create a new scheme"
subtitle: "`r library(dmdScheme);paste0(scheme_active()$name, '_', scheme_active()$version)`"
author: "Rainer M Krug <Rainer.Krug@uzh.ch>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Howto_create_new_scheme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This package contains all the functionality to easily create a new metadata scheme. The resulting scheme package will contain all files needed so that it can be easily be distributed and used, and can be either be distributed directly or uploaded to an online repository, e.g. the default [dmdScheme repository](https://github.com/Exp-Micro-Ecol-Hub/dmdSchemeRepository).

# Create new scheme

To create your own scheme package entails a few simple steps.

1. Open an existing scheme definition you want to use as a starting point. The easies is to use the command   

```{r eval=FALSE, echo=TRUE}
open_new_spreadsheet(keepData = TRUE)
```

which will open the current scheme including example data in your spreadsheet program of choice.

**If the `.xlsx` file is locked, you can unlock it by using the password `test`.**

You can also manually extract the `.xlsx` file from any scheme package and use it.

It is important, that the final scheme contains example data as the resulting scheme package has to contain example data.

2. Edit the new scheme definition

This `.xlsx` file can be edited following the limitations as specified in the **Minimum requirements for metadatascheme based on dmdScheme**. After modifications have been done and the spreadsheet has been saved, the scheme package can be created by using

```{r eval = FALSE}
scheme_make( "TheFile.xlsx" )
```

which will create a file named `NAME_VERSION.tar.gz` in the directory `path` based on the scheme definition in `TheFile.xlsx`. `NAME` and `VERSION` are the scheme name and scheme version as specified in the cell `(H:1)` in sheet `Experiment`.

The following two commands will install the new scheme and use it:

```{r eval = FALSE, echo = TRUE}
scheme_install(file = "NAME_VERSION.tar.gz")
scheme_use("NAME", "VERSION")
```

To upload the scheme to the main repository on [github](https://github.com/Exp-Micro-Ecol-Hub/dmdSchemeRepository), please either clone the repo and send a pull request or file an issue with the scheme package as an attachment.

# Create R package to Extending the functionality of the `dmdScheme` package
The package `dmdScheme` package is only providing the base functionality for working with dmdSchemes. In many cases, the functionality needs to be extended to be able to work with other schemes. This can be done by creating a new package which is

1. depends on the package `dmdScheme`, and
2. automatically downloads, installs and activates a specific scheme definition and not the default dmdScheme.

Similar to the function `package.skeleton()` in the `utils` package, `dmdScheme` provides a function which creates a package skeleton with the required functionality. 

To for example create a package in the folder `path = tempdir()` which depends on the R package `dmdScheme` and loads and activates upon loading the scheme `emeScheme` of version `0.9.5` which is in the default repository:
 
```{r eval = TRUE, echo = TRUE}
if (!scheme_installed("emeScheme", "0.9.5")) {
  scheme_install("emeScheme", "0.9.5")
}
scheme_use("emeScheme", "0.9.5")

make_new_package( path = tempdir() )

scheme_use("dmdScheme", "0.9.5")
scheme_uninstall("emeScheme", "0.9.5")
```

**NOTE: If the scheme package is not in the default repository, the file `inst/aaa.R` in the new package needs to be adapted to automatically anstall and use the scheme!**

For an example of an R package of this kind see the R folder in the [emeScheme package on github](https://github.com/Exp-Micro-Ecol-Hub/emeScheme).
