---
title: "User manual for experimental microbial ecology metadata tool"
author: "Owen and Rainer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

Researchers are increasingly being required to make the data they produce findable, accessible, interoperable, and reusable [wikipedia FAIR entry](https://en.wikipedia.org/wiki/FAIR_data). Providing metadata that describes our studies, experiments, treatments, measurements, and so on is a key part of this endeavour.

The aim of the experimental microbial ecology metadata tool is to help you describe your study and its data, i.e. to produce useful metadata. You will do this by preparing your datasets for archiving (more on this below), filling in your metadata the spreadsheet supplied, validating the combination of the metadata and data (this will be largely automatic), fixing any issues revealed by the validation step, and once satisfactorily validated deposit the metadata and data in a repository of your choice.

What is meant by *useful* metadata? Imagine we would like to use the metadata to find experiments (and potentially return the relevant data from them) that involve *Colpidium striatum* grown at 20ºC in 0.55g per litre protist pellet medium. To do this we need to ensure that the species name, the temperature, and the growth media information are contained in the metadata. We also need to make sure that when temperature is described it is described in a standard fashion, and the the growth media description is standardised. So there must be some standardisation. But its clear that most experiments have some unique features, such that we can't standardise everything in advance. The metadata scheme we use here is intended to be a good balance of standardised and flexible.

# When should I enter metadata

You will benefit from starting this when you design your study. You'll likely find out that doing to will help with the planning of the study. You may then choose to work in a slightly different way, since this will make things more efficient later (this might especially be so for data handling issues). No problem, however, if you decide to do it at the end (or you are reading this after a study is complete).

# Getting started

You will enter the metadata in a supplied spreadsheet (Googlesheets or Excel), but then use R to do some tasks, like validation and transormation of metadata into more machine readable form. So the first thing to do is get and load the **emeScheme** R package:

```{r}
## install the devtools package if not installed yet
## install.packages("devtools")
## install the last version of emeScheme from github
#devtools::install_github("Exp-Micro-Ecol-Hub/emeScheme")
library(emeScheme)
```

To get a copy of the spreadsheet you will enter the metadata in, you can run this R command to get an Excel version, or use [this copy on googlesheets](https://docs.google.com/spreadsheets/d/1sCImO4VadNwNKuqgtC3CPksS95Tl6SJVtY78aFo9gLg): 

```{r eval=FALSE}
enter_new_metadata()
```

Either way, make a copy of the spreadsheet and save it somewhere safe, like in the folder for your study/experiment.

Its probably worth now looking at the spreadsheet (but don't yet try to fill in anything), to see what metadata you will need to enter. There are several worksheets:

* **Experiment** Here you enter details of your experiment, e.g. what temperature was the experiment conducted at. If temperature was a treatment, then you state that here and supply the levels in the **Treatments** worksheet.
* **Species** Give details of the species involved in the experiment.
* **Treatments** Here we give the manipulations involved in the experiment.
* **Measurements** Details of what was measured, and how data was extracted from the measurements.
* **DataExtraction** Details of how measurements were extracted from raw data, if that happened.
* **DataFileMetaDat** Details of the data files, description of the variables they contain, and validation conditions for the data.

# Preparating data for archiving

There are no hard and fast rules, but there are some guidelines that will make things easier, or are otherwise recommended for datasets (e.g. data that we can have in a spreadsheet):

* Use long / tidy arrangement of data.
* Always include a variable for each treatment.
* Always include a variable that allows one data file to be related to others (e.g. Microcosm_ID).
* Put all measurement of one type in a single data file. E.g. one datafile of dissolved oxygen measures, one datafile of adundance data, one data file of sequence data, containing all measurements from all treatments.

If your data does not fit these guidelines, e.g. is a collection of videos files or a collection of files containing DNA sequence data, put it in a container of some kind (e.g. compressed folder or zip file) and describe this in the metadata.

## A worked example of preparation of data for archiving

The worked example shows appropriate preparation of data for archiving and ease of further use. Its an example of what you could do, not what you have to do, though probably you benefit from doing something similar.

Get the worked example folder and files therein from (here)[https://drive.google.com/open?id=14ciYlIvQZ8DehMfqDE7s5K5GlYbZpqsH]. Put the folder somewhere on your computer and have a look in it. You will see:

* An R-project file `expt1.Rproj`
* A googlesheet file `expt1 emeScheme.gsheet`
* A folder named `code` and a folder named `data`
* The `data` folder contains two folders, one named `raw_data` and one named `archiving_data`.
* The `raw_data` folder contains files and folders of files that contain the data as it was entered by the researcher, or produced by the measurement machine.
* The `archiving_data` folder contains data files that have been prepared for archiving.
* The `code` folder contains two files: `data_preparation.Rmd` and `analyses.Rmd`.
* `data_preparation.Rmd` contains code that takes the information in the files in the `raw_data` folder, prepares this information for easy use by researchers (which also makes them suitable for archiving), and then writes this nicely prepared information/data into files in the `archiving_data` folder.
* `analyses.Rmd` starts by reading the nicely prepared data files from the `archiving_data` folder, and would then proceed with analyses.

**You will not be able to use the code in `data_preparation.Rmd` as it is highly specific to this example.**

**Note** that it is **very important** thet the preparation of the data for easy use by researchers and for archiving (done in `data_preparation.Rmd`) in no way changes, subsets, or otherwise manipulates the data. It is just a rearrangement of the raw data and addition of treatment and other useful information.

Super. Now we have nicely prepared data. We are ready to fill in the metadata.

# Fill in the metadata

There's no more putting it off... lets try filling in the metadata. First find and open your copy of the metadata spreadsheet you made.

In the worked example, open the `expt1 emeScheme.gsheet` file.

* Green cells are where you enter information.
* Pink cells are places where you should not try to type anything (they should be locked anyway).

## Experiment worksheet

This one should not need too much explanation. One important point is, however, what to enter in a field (e.g. temperature) if your experiment involved different temperatures (i.e. temperature was a treatment). The answer is you should enter the word "treatment".

When you fill in this information, think about standardisation. If you enter a temperature, you should therefore enter only a numeric value in degrees Celcius. Writing "twelve degrees" is going to be quite unhelpful! (We could constrain the possible entries here, but then we lose flexibility; recall the discussion above about balancing standardisation and flexibility.)

Something else of note: Look at the field "media Additions" and you will see the entry is "Wheat seeds added on specific dates, see file wheat_seed_additions.csv". Here we see the possibility for great flexibility... we refer to information in a separate data file. Note though that this data file is just a list of dates and number of wheat seeds added--it is not information about an experimental treatment, because all experimental treatments experience the same additions (i.e. these additions are feature of the experiment).

## Species worksheet

This worksheet should also not need much explanation. Note that this is just a list of the species in the experiment. Its not a description of any treatments involving species composition.

During validation of your metadata the species names will be automatically checked against an online database of species names, and you will get a report on matches or lack of matches. (This helps increase interoperability of the data.)

In the worked example the experiment was set up by adding a pool of unknown species composition. So no entries in this worksheet. The composition could be "measured" using e.g. molecular methods, and would then be an entry in the **Measurements** worksheet (see below for more details on that worksheet).

If any of your data sets contain species names, it will be very helpful to make them match the names in this sheet. You will be told about the presence or absence of such matches in the validation report.

## Treatments worksheet

Here enter in each row each of the treatment levels in each treatment. In the "parameter" column put the name of the treatment (in the worked example this is "Lid_treatment").

You do not need to, and should not, attempt to include information about treatment combinations. So if your experiment involved a temperature treatment with three levels (15, 20, and 25) and a species composition treatment with four treatments (species A alone, species B alone, species C alone, and all three together) you would enter seven rows of information. You would not enter 12 rows if your experiment was a two-way fully factorial design with these two treatments.

It would be very nice here to make the names of your treatment and the descriptions of the levels exactly the same as those used in your data files. You will be told about the presence or absence of such matches in the validation report.

## Measurements worksheet

Here enter a row of inforamtion for each type of measurement made. In the worked example there are four types of measurement: oxygen concentration, community composition, smell, and DNA sequences. Much of the information required in the rows should be quite self explanatory, apart perhaps from:

* Data extraction method: Fill in something here if the measurement required some processing of another data file / measurement. E.g. in the worked example the commmunity composition measure comes from the DNA data by extraction method "Mol_Analy_pipeline1".
* Source measurement: should be one of the names of the other rows of measurements. E.g. in the worked example the community composition measure comes from the "sequenceData" measurement.

It would be very nice/useful if the entries in the "variable" column correspond with variables in the data file mentioned in the "data file" column. You will be told about the presence or absence of such matches in the validation report.

## DataExtraction

Here you can give details of any (and please all - this will be validated in the validation report) data extraction methods mentioned in the **Measurements** worksheet. You can enter any parameters and values of these you like (i.e. there is high flexibility here because we do not anticipate a lot of standardisation to exploit). In the worked example we wrote "See description in file xxx.yyy". This file could be a script or text description of the pipeline.

## DataFileMetaData worksheet

Here you give a row for each of the variables in each of the data files.

* **dataFileName** Here enter the file name of the data file.
* **columnname** Here enter the exact variable name used in the datasheet / database / tabular data.
* **measurementName** Here enter the measurement name, corresponding to the exact measurement name used in the **Measurement** worksheet, if the column contains a measurement.
* **unit** Here enter the measurement units (e.g. % or grams).
* **type** Enter the variable type (e.g. numeric, character, date).
* **description** Anything you think relevant.
* **treatmentParameter** Here enter the treatment name corresponding to the exact treamtent name used in the **Treatment** worksheet, if the column contains treatment information.

You will get a report on things such as matches and lack of matches with entered variable names and those in the actual datasets, matches of treatment names and measurement names, absences of data about things that were measured, absence about treatment information for things that were measured, and so on.

## Validation

**<span style="color:red">To be done..</span>**.


