---
params:
  author: unknown
  title: Validation
  x: NULL
  result: NULL
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
output:
  html_document: 
    number_sections: true
    toc_depth: 3
    toc: true
    toc_float: true
    keep_md: true
  pdf_document: 
    number_sections: true
    toc_depth: 3
    toc: true
  word_document: 
    number_sections: true
    toc_depth: 3
    toc: true
---


```{r setup, include=FALSE}
library(magrittr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
x <- params$x
result <- params$result

## only for debugging ##
if (is.null(x)) {
  if (!require("emeScheme")) {
    devtools::load_all()
  }
  x <- emeScheme::emeScheme_raw
  result <- validate(x = emeScheme_raw, validateData = TRUE, report = "none", path = system.file("example_data", "basic", "data", "archiving_data", package = "emeScheme"))
}
## ##
## Define colours for results of evaluation

```

* **Author**: `r params$author`
* **emeScheme metadata name**: `r attr(x, "fileName")`
* **data package path**: `add this` 

# Introduction
This validation report validates the most likely an excelsheet containing the data.
The headers do represent sections of validations, while the bullets reperesent individual validations. Sections of the validation and are hierarchical, i.e. a higher level validation result depends on the lower level validation results. The headers and validations follow the structure of

**Error_Level** : **Name of the validation**

The error level can be:
- **OK**: the validation passed
- **note**: the validation contains an inconsistency which migh be an error or on purpose. Please check these!
- **warning**: the validation contains an inconsistency which could be an error or on purpose. Please check these!
- **error**: the validation failed and contains an error - needs to be fixewd before the data can be exported!

An ideal validation will not contain any errors or warnings.

Validation results are accumulated by forwarding the highest error to the validation of the next 
level. If validation 1.3. contains an error, 1. will be classified as an error as well, 
irrespective of 1.1, 1.2, ....

The details do contain the individual tests and are explained in detail in the sections

Details **Error_Level** : **Name of the validation**


<!-- ################################### -->
<!-- ################################### -->
<!-- Overall MetaData -->
<!-- ################################### -->
<!-- ################################### -->

# Overall
```{r}
valErr_extract(result) %>% 
  table %>% 
  set_names(valErr_info(names(.))$text)
```

# Overview
```{r, results = "asis"}
print(result, level = 2, listLevel = 20, type = "summary")
```

## **<span style="color:red">TODO</span>** 
### result$DataFiles$header
result$DataFiles$description

- [ ] ranges of numeric columns
- [ ] values of text columns
- [ ] ???

### <== Treatment
- [ ] the **values** in the column describing the treatment (i.e. treatment levels) have to be in the column `Treatment$treatmentLevel` of the corresponding `Treatment$parameter`

### <== Measurement
- [ ] type checks of `columnName` compared to `type`

## Other Validity Checks
- [ ] ???


<!-- ################################### -->
<!-- ################################### -->
<!-- Details -->
<!-- ################################### -->
<!-- ################################### -->


# Details

```{r , results = "asis"}
print(result, level = 2, listLevel = 20, type = "details", format = "markdown")
```



# Structure info
```{r info}
attributes(x)
```

# Confirmation code

The following code needs to be entered in the `emeSchemeToXml()` function to confirm that the validation has been done and read. **No export is possible without this code!**:
 **<span style="color:red">TODO implement in emeSchemeToXML()</span>**
```{r echo=FALSE, comment = ""}
if (isTRUE(result$error < 2)) {
  suppressWarnings( 
    new_emeSchemeSet(
      x = x, 
      keepData = TRUE, 
      convertTypes = TRUE,  
      verbose = FALSE, 
      warnToError = FALSE
    ) 
  ) %>%
    digest::digest(algo = "sha1")
} else {
  cat_ln("Please fix all errors before continuing!")
  cat_ln()
  cat_ln("The following code will not be printed in the final version!!!!!")
  cat_ln()
  suppressWarnings( 
    new_emeSchemeSet(
      x = x, 
      keepData = TRUE, 
      convertTypes = TRUE,  
      verbose = FALSE, 
      warnToError = FALSE
    ) 
  ) %>%
    digest::digest(algo = "sha1") %>%
    cat_ln()
}
```


# <span style="color:red">TODO</span>  
- Use conditional text to explicitly say pass or fail, so that users don't have to read the R output (or at least get lovely green text for pass, and red for fail!)
- Validation out of attempt to parse date variables
