---
title: "Experimental Microbial Ecology Metadata Scheme"
author: "Rainer M. Krug"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
  pdf_document: default
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
nocite: '@*'
bibliography: emeScheme_papers.bib
link-citations: TRUE
---

```{r setup, echo = FALSE, include = FALSE}
library(here)
library(magrittr)
library(knitr)
library(kableExtra)
library(plantuml)
##
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 6,
	fig.width = 6,
	comment = NA,
	cache = FALSE
)
##
devtools::load_all(here())
```

# Introduction

This scheme will provide a framework for describing aquatic microcosm experiments in the sense as used in the papers in the [Reference] section.

This scheme **aims to** cover the following aspects:

1. **Materials**: these include instrumentation, species, cultures, ...
1. **Initialization**: run in time, treatment, conditions
2. **Conditions / Manipulations**: different conditions for the different treatments and manipulations, conditions, transplantations, ...
3. **Sampling** (how much is taken from microcosms and how much is re-filled with what and how often, ...)
3. **Measuring** of the raw data (video, images, count, ...) using different methodologies and their instrumentations, protocols, ...
4. **Extraction** of the needed measure  (density, ...) from the measured raw data (if needed) and the software, versions, parameter used (e.b. bemovi)

By using this metadata one should be able to successfully re-generate the data.

This scheme **does not** aim to cover aspects of the actual analysis of this data with the aim of answering specific questions.

# The emeScheme

## Experimental Microbial Ecology Protocols
In the  [supplement to their paper](https://emeh-protocols.readthedocs.io/en/latest/) to their paper, @altermatt_big_2015 specify the following sections for a protocoll. The sections are as followed:

```
1. Materials
1.1 Species used
1.2 Culture medium
1.3 Bacteria
1.4 Apparatus
1.5 Lab practices
1.6 Long term maintenance
1.7 Long term preservation
2. Measurements
2.1 Sampling
2.2 Microscopy
2.3 Image analysis
2.4 Particle counter
2.5 Bacterial density
2.6 Raman microspectroscopy
2.7 Barcoding
2.8 Ge-, prote- and epigenomics
2.9 Respirometry
2.10 Nutrients and decomposition
2.11 Protocol unavailable
2.12 Interactions
3. Manipulations
3.1 Protocol unavailable
3.2 Density
3.3 Disturbance
3.4 Nutrients and viscosity
3.5 Spatial structure
3.6 temperature
3.7 Biotic environment
```

We use these sections as a starting point to define a metadata scheme to describe the microcosm experiment. 


## Supplementing, Re-ordfering and Thinning

The sections in the [Experimental Microbial Ecology Protocols] were supplemented and assessed by reviewing the methods in the papers in the [References] section below. This resulted in the document [emeScheme.numbers].

This document was than re-ordered and thinned which resulted in the initial verions of [the Google Doc Sheet](https://docs.google.com/spreadsheets/d/1OAyRM1jGL5Vho-YfPffePhwGpQtaBUeTPJV6P98f0Bc).

Here we describe the different properties of the emeScheme We will start at the first level and go deeper into the different properties.

The working copy can be found on a [Google Doc Sheet](https://docs.google.com/spreadsheets/d/1OAyRM1jGL5Vho-YfPffePhwGpQtaBUeTPJV6P98f0Bc). 



## Current Structure

The names of the proiperties folow the following convention:

* Properties starting with a **capital letters** contain sub-properties, i.e. can be considered as **property sets**
* Properties starting with a **small letter** contain actual values of the types **character** or **numeric**. **<span style="color:red">types need to be specified</span>**

**<span style="color:red">Inclusde mandatory, repeatable, ...</span>**
The structure as currently (last version updated at `r read.dcf(here("DESCRIPTION"))[1, "GSUpdate"]`) is as followed:

```{r structure}
caption <- "emeScheme Properties. "

options(knitr.kable.NA = '')

emeScheme %>% 
  knitr::kable(
    format = "html",
    caption = caption
  )  %>% 
  kable_styling(
    bootstrap_options = c("striped", "bordered"), 
    full_width = FALSE
  )
```

# From Project to Dataset - how the emeScheme fits in

Data can be separated into different classes. Each of these classes sontains of **one** (data) file and **one** metadata file. 

These metadata files contain in the property **RelatedIdentifier** providing the possibility to link different (meta)data sets. 


```{r data_layout, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE, fig.width = 33, fig.height = 49}
'
class Project {
  {field} Project description
  {method} - metadata
}

class Experiment_1 {
  {field} Experimental description
  {method} + Experimental Layout
}
class Experiment_n {
  {field} Experimental description
  {method} + Experimental Layout
}


class Treatment_1 {
  {field} Treatment Description
  {method} + Materials
  {method} + Initialization
  {method} + Conditions
}

class Treatment_n {
  {field} Treatment Description
  {method} + Materials
  {method} + Initialization
  {method} + Conditions
}

class Measurement_1 {
  {field} Measurement description
  {method} + Sampling
  {method} + Measuring and Extraction
}

class Measurement_n {
  {field} Measurement description
  {method} + Sampling
  {method} + Measuring and Extraction
}

class RawData {
  {field} Measured Data
  {method} - metadata
}

class ExtractedData_1 {
  {field} Extracted Data
  {method} - metadata
  {method} + Measuring and Extraction
}

class ExtractedData_n {
  {field} Extracted Data
  {method} - metadata
  {method} + Measuring and Extraction
}


Project "1" *-- "many" Experiment_1 : contains
Project "1" *-- "many" Experiment_n : contains

Experiment_n "1" *-- "many" Treatment_1 : contains
Experiment_n "1" *-- "many" Treatment_n : contains

Treatment_n "1" *-- "many" Measurement_1 : contains
Treatment_n "1" *-- "many" Measurement_n : contains
Measurement_n "1" *-- "1" RawData : contains
RawData "1" *-- "many" ExtractedData_1 : contains
RawData "1" *-- "many" ExtractedData_n : contains
' %>%
  plantuml %>%
  plot()
```



# References

