---
title: "Outline Metadata"
author: "Rainer M. Krug"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
  pdf_document: default
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
nocite: '@*'
bibliography: microcosm_papers.bib
link-citations: TRUE
---

```{r setup, echo = FALSE, include = FALSE}
library(here)
library(magrittr)
library(devtools)
library(googlesheets)
library(dplyr)
# library(magrittr)
library(knitr)
library(kableExtra)
##
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 6,
	fig.width = 6,
	comment = NA,
	cache = FALSE
)
##
gs_auth(token = "~/keys/googlesheets_token.rds")
mcs <- gs_title("microcosmScheme")
microcosmScheme_gd <- 
   mcs %>%
  gs_read()
gs_deauth()
microcosmScheme <- microcosmScheme_gd %>%
  select(starts_with("Property"))

notNA <- microcosmScheme %>% 
  is.na() %>% 
  rowSums() %>% 
  equals(ncol(microcosmScheme)) %>%
  not()
microcosmScheme <- microcosmScheme[notNA,]
  
save( microcosmScheme_gd, file = here("data", "mS_gd.rda"))
pkgPath <- here("data")
if (
  tools::md5sum(here("data", "microcosmScheme_gd.rda")) == tools::md5sum(here("data", "mS_gd.rda"))
) {      ##### no change in google sheet
  unlink( file.path(pkgPath, "mS_gd.rda") )
} else { ##### change in google sheet
  unlink( here("data", "mS_gd.rda") )
  unlink( fhere("data", "microcosmScheme_gd.rda") )
  save( microcosmScheme_gd, file = here("data", "microcosmScheme_gd.rda"))
  unlink( here("data", "microcosmScheme.rda") )
  save( microcosmScheme, file = here("data", "microcosmScheme.rda"))
  ##### bump version #####
  ## read old DESCRIPION file
  DESCRIPTION <- read.dcf(here("DESCRIPTION"))
  ##
  ## Increase Version
  currVersion <- DESCRIPTION[1, "Version"]
  splitVersion <- strsplit(currVersion, ".", fixed = TRUE)[[1]]
  nVer <- length(splitVersion)
  currEndVersion <- as.integer(splitVersion[nVer])
  newEndVersion <- as.character(currEndVersion + 1)
  splitVersion[nVer] <- newEndVersion
  newVersion <- paste(splitVersion, collapse = ".")
  DESCRIPTION[1, "Version"] <- newVersion
  ##
  ## Set Date
  DESCRIPTION[1, "Date"] <- format(Sys.time(), '%Y-%m-%d')
  ##
  ## adjust Description
  DESCRIPTION[1, "Description"] %<>%
    strsplit(., "\n----\n") %>% 
    extract2(1) %>%
    extract(1) %>%
    paste0("\n----\n", "Data imported and version bumped at ", date(), " / ", Sys.timezone())
  
  ## write new DESCRIPTION
  write.dcf(DESCRIPTION, here("DESCRIPTION"))
  rm( microcosmScheme, microcosmScheme_gd )
  devtools::load_all(here())
}
```

# Introduction

This scheme will provide a framework for describing aquatic microcosm experiments in the sense as used in the papers in the [Reference] section.

This scheme **aims to** cover the following aspects:

1. **Materials**: these include instrumentation, species, cultures, ...
1. **Initialization**: run in time, treatment, conditions
2. **Conditions / Manipulations**: different conditions for the different treatments and manipulations, conditions, transplantations, ...
3. **Sampling** (how much is taken from microcosms and how much is re-filled with what and how often, ...)
3. **Measuring** of the raw data (video, images, count, ...) using different methodologies and their instrumentations, protocols, ...
4. **Extraction** of the needed measure  (density, ...) from the measured raw data (if needed) and the software, versions, parameter used (e.b. bemovi)

By using this metadata one should be able to successfully re-generate the data.

This scheme **does not** aim to cover aspects of the actual analysis of this data with the aim of answering specific questions.

# The microcosmScheme

## Experimental Microbial Ecology Protocols
In the  [supplement to their paper](https://emeh-protocols.readthedocs.io/en/latest/) to their paper, @altermatt_big_2015 specify the following sections for a protocoll. The sections are as followed:

```
1. Materials
1.1 Species used
1.2 Culture medium
1.3 Bacteria
1.4 Apparatus
1.5 Lab practices
1.6 Long term maintenance
1.7 Long term preservation
2. Measurements
2.1 Sampling
2.2 Microscopy
2.3 Image analysis
2.4 Particle counter
2.5 Bacterial density
2.6 Raman microspectroscopy
2.7 Barcoding
2.8 Ge-, prote- and epigenomics
2.9 Respirometry
2.10 Nutrients and decomposition
2.11 Protocol unavailable
2.12 Interactions
3. Manipulations
3.1 Protocol unavailable
3.2 Density
3.3 Disturbance
3.4 Nutrients and viscosity
3.5 Spatial structure
3.6 temperature
3.7 Biotic environment
```

We use these sections as a starting point to define a metadata scheme to describe the microcosm experiment. 


## Supplementing, Re-ordfering and Thinning

The sections in the [Experimental Microbial Ecology Protocols] were supplemented and assessed by reviewing the methods in the papers in the [References] section below. This resulted in the document [microcosmScheme.numbers].

This document was than re-ordered and thinned which resulted in the initial verions of [the Google Doc Sheet](https://docs.google.com/spreadsheets/d/1OAyRM1jGL5Vho-YfPffePhwGpQtaBUeTPJV6P98f0Bc).

Here we describe the different properties of the microcosmScheme We will start at the first level and go deeper into the different properties.

The working copy can be found on a [Google Doc Sheet](https://docs.google.com/spreadsheets/d/1OAyRM1jGL5Vho-YfPffePhwGpQtaBUeTPJV6P98f0Bc). 



## Current Structure

The names of the proiperties folow the following convention:

* Properties starting with a **capital letters** contain sub-properties, i.e. can be considered as **property sets**
* Properties starting with a **small letter** contain actual values of the types **character** or **numeric**. **<span style="color:red">types need to be specified</span>**

**<span style="color:red">Inclusde mandatory, repeatable, ...</span>**
The structure as currently (last version updated at `r mcs$updated`) is as followed:

```{r structure}
caption <- "microcosmScheme Properties. "

options(knitr.kable.NA = '')

microcosmScheme %>%
  select(starts_with("Property")) %>% 
  knitr::kable(
    format = "html",
    caption = caption
  )  %>% 
  kable_styling(
    bootstrap_options = c("striped", "bordered"), 
    full_width = FALSE
  )
```



# References

