---
title: "Outline Metadata"
author: "Rainer M. Krug"
date: "11/7/2018"
output:
  html_document: default
  pdf_document: default
nocite: '@*'
bibliography: microcosm_papers.bib
link-citations: TRUE
---

```{r setup, echo = FALSE}
library(magrittr)
library(knitr)
library(kableExtra)
##
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 6,
	fig.width = 6,
	comment = NA
)
```

# Introduction

This scheme will provide a framework for describing aquatic microcosm experiments in the sense as used in the papers in the [Reference] section.

This scheme **aims to** cover the following aspects:

1. **Materials**: these include instrumentation, species, cultures, ...
1. **Initialization**: run in time, treatment, conditions
2. **Conditions / Manipulations**: different conditions for the different treatments and manipulations, conditions, transplantations, ...
3. **Sampling** (how much is taken from microcosms and how much is re-filled with what and how often, ...)
3. **Measuring** of the raw data (video, images, count, ...) using different methodologies and their instrumentations, protocols, ...
4. **Extraction** of the needed measure  (density, ...) from the measured raw data (if needed) and the software, versions, parameter used (e.b. bemovi)

By using this metadata one should be able to successfully re-generate the data.

This scheme **does not** aim to cover aspects of the actual analysis of this data with the aim of answering specific questions.

# Experimental Microbial Ecology Protocols
In the  [supplement to their paper](https://emeh-protocols.readthedocs.io/en/latest/) to their paper, @altermatt_big_2015 specify the following sections for a protocoll. The sections are as followed:

```
1. Materials
1.1 Species used
1.2 Culture medium
1.3 Bacteria
1.4 Apparatus
1.5 Lab practices
1.6 Long term maintenance
1.7 Long term preservation
2. Measurements
2.1 Sampling
2.2 Microscopy
2.3 Image analysis
2.4 Particle counter
2.5 Bacterial density
2.6 Raman microspectroscopy
2.7 Barcoding
2.8 Ge-, prote- and epigenomics
2.9 Respirometry
2.10 Nutrients and decomposition
2.11 Protocol unavailable
2.12 Interactions
3. Manipulations
3.1 Protocol unavailable
3.2 Density
3.3 Disturbance
3.4 Nutrients and viscosity
3.5 Spatial structure
3.6 temperature
3.7 Biotic environment
```

We use these sections as a starting point to define a metadata scheme to describe the microcosm experiment. The sections are supplemented byn evaluating the papers in the [References] section below.

# The microcosmScheme

**<span style="color:red">Read from GoogleSheets and save to csv. This can tham be annotated here.</span>**

Here we describe the different properties of the microcosmScheme We will start at the first level and go deeper into the different properties.

## First Level Propertiers

```{r prop_l1}
caption <- "microcosmScheme first level Properties"

p1 <- data.frame(
  ID = 1:5,
  Property = c("Materials", "Initialization", "Conditions", "Sampling", "MeasuringAndExtraction"),
  Description = c(
    "instrumentation, species, cultures, ..", 
    "run-in time, conditions, densities, ...",
    "for all treatments conditions, and manipulations during experiment", 
    "how the sample is obtained", 
    "measuring and extracting the measures from the sample"
  )
)


knitr::kable(
  p1, 
  format = "html",
  caption = caption
)  %>% 
  kable_styling(
    bootstrap_options = c("striped", "bordered"), 
    full_width = FALSE
  )
```



# References

